 

### 2021-07-01 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

$Ga$ and $Gs$ are dominant factors in $T_{leaf}$ flux. $Ga$ increases with wind speed which also in turn increases $gs$, thereby reduces $T_{leaf}$-$T_{air}$ (Fig. `r fig_leaf_T`b). Therefore, high windspeeds in upper canopies such as in NEON sites with windspeeds ~4 ms (Fig. `r fig_NEON_vertical`) are capable of lowering $T_{leaf}-T_{air}$ that are exposed to high radiation loads, upto ~ 4$^\circ$C (Fig. `r fig_leaf_T`b)[Bonan 2008]. However, low wind speeds typically decouple leaves from $T_{air}$, more so when solar radiation is high and $gs$ is low.

@@

**Within forests, vertical gradients in the propensity for $T_{leaf}$ to be elevated above $T_{air}$ are expected to vary with micrometeorlogical conditions (Fig. `r fig_leaf_T`).**
$T_{leaf}$ generally tracks $T_{air}$, and in turn the vertical gradient of $T_{air}$ depends on canopy structure and $LAI$. Closed canopies, and high $LAI$ act as a sheild, absorbing most of the incoming radiation and preventing vertical air mixing in the understory microclimate. Therefore, we would expect a stronger $T_{leaf}$ gradient where upper canopy leaves can be much warmer than below canopies layers [@millerOnlySunlitLeaves2021]. On the other hand, open canopies, or lower canopy $LAI$ allow more vertical air mixing and sunlight into below canopies, that can either neutralize $T_{leaf}$ gradient or in the case of 'canopy greenhouse effect' or even sunflecks, can elevate below canopy $T_{leaf}$-$T_{air}$ of shade leaves more than upper canopy due to the propensity of shade leaves mentioned above [@hardwickRelationshipLeafArea2015; @schymanskiStomatalControlLeaf2013]. 
It is also important to note that NEON sites (in Fig. `r fig_NEON_vertical`) might not exactly reflect the $T_{leaf}$ pattern we might expect, partly because $T_{bio}$, measured using infrared sensor, reflects wide swath of temperatures. But recent advances in $T_{leaf}$ research that we highlight in the Modelling section will enhance our understanding of large scale $T_leaf$ vertical patterns.~

>>>>

$Ga$ and $Gs$ are dominant factors in $T_{leaf}$ flux. $Ga$ increases with wind speed which also in turn increases $gs$, thereby reduces $T_{leaf}$-$T_{air}$ (Fig. `r fig_leaf_T`b). Therefore, high windspeeds in upper canopies such as in NEON sites with windspeeds ~4 ms (Fig. `r fig_NEON_vertical`) are capable of lowering $T_{leaf}-T_{air}$ that are exposed to high radiation loads, upto ~ 4$^\circ$C (Fig. `r fig_leaf_T`b)[Bonan 2008]. However, low wind speeds typically decouple leaves from $T_{air}$, more so when solar radiation is high and $gs$ is low. Additionally, $T_{leaf}-T_{air}$ increases with humidity (Fig. `r fig_leaf_T`c) because reduced $ga$ stifles heat dissapation through $E$ and $gs$.

@@

**Within forests, vertical gradients in the propensity for $T_{leaf}$ to be elevated above $T_{air}$ are expected to vary with micrometeorlogical conditions (Fig. `r fig_leaf_T`).**
$T_{leaf}$ generally tracks $T_{air}$, and in turn the vertical gradient of $T_{air}$ depends on canopy structure and $LAI$. Closed canopies, and high $LAI$ act as a sheild, absorbing most of the incoming radiation and preventing vertical air mixing in the understory microclimate. Therefore, we would expect a stronger $T_{leaf}$ gradient where upper canopy leaves can be much warmer than below canopies layers [@millerOnlySunlitLeaves2021]. On the other hand, open canopies, or lower canopy $LAI$ allow more vertical air mixing and sunlight into below canopies, that can either neutralize $T_{leaf}$ gradient or in the case of 'canopy greenhouse effect' or even sunflecks, can elevate below canopy $T_{leaf}$-$T_{air}$ of shade leaves more than upper canopy due to the propensity of shade leaves mentioned above [@hardwickRelationshipLeafArea2015; @schymanskiStomatalControlLeaf2013]. 
It is also important to also note though NEON sites (in Fig. `r fig_NEON_vertical`) might not exactly reflect the $T_{leaf}$ pattern we might expect, partly because $T_{bio}$ measured using infrared sensor, reflects a wide swath of biological temperatures, but recent advances in $T_{leaf}$ modelling [see Scaling across Space and Time] will enhance our understanding of large scale $T_leaf$ vertical patterns.~~
